Designing Types for R, Empirically

### 2020-05-16 paper/main.tex
edits

\typetracer observed that \PROPSTRUCTSDYN of lists had a \code{names} attribute, and in R, ``fields'' of named lists can be accessed using the
\code{$} operator.

\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
  agricolae::cv.model <- function(x) {
    suma2 <- sum(x$residual^2)
    gl <- x$df.residual
    promedio <- mean(x$fitted.values)
    return(sqrt(suma2/gl)*100/promedio)
  }
  data(sweetpotato)
  model<-aov(yield~virus, data=sweetpotato)
  cv.model(model)
\end{lstlisting}

In the code snippet, the function \code{cv.model} takes an argument
\code{x}, which we observed to always be a {\it linear model}, which is part
of R's built-in statistics functionality.  Internally, linear models are
represented as lists with named elements, and one may access these named
elements using the \code{$} syntax, as seen on lines 2-4 in the snippet.  To
describe these kinds of values, we included a struct type in our type
system, allowing users to specify name, type pairs as part of list type
definitions.

Unfortunately, the inclusion of structs complicated our analysis by
polluting the data, illustrated on line 8 with the call
\code{data(sweetpotato)}.  It is quite common for R example code to use some
built-in data sets, such as \code{sweetpotato}, to show off some package
functionality with ``real data''.  When \typetracer is configured to detect
structs we saw a marked increase in misrepresentative types, as the types of
functions which use these data sets (such as \code{aov} in
Figure~\ref{fig:struct-ex}) became polluted with the names from the example
sets.  We implemented a unification strategy for struct types, removing them
if they co-occurred with list types, and turning them into lists if two
co-occurring struct types shared no common names, but even so we found
ourselves with function signatures that were full of noise introduced by the
struct names.

We still believe struct types to be useful to R programmers, and so our
language supports the type and \contractr will check them, though we
configured \typetracer to ignore them as they made the data substantially
less useful.  On the positive side, we found that \AT{get the number, but
  often} structs-typed values had a {\it class}, which helped motivate our
inclusion of class-like types in our type annotation language.

>>>>

In R, lists can have a names attribute, allowing lists elements to be accessed by name using the
\code{$} operator.

\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
#' @type <x: class<aov, lm> | class<lm>> => dbl;
agricolae::cv.model <- function(x) {
    suma2 <- sum(x$residual^2)
    gl <- x$df.residual
    promedio <- mean(x$fitted.values)
    return(sqrt(suma2/gl)*100/promedio)
  }
  data(sweetpotato)
  model<-aov(yield~virus, data=sweetpotato)
  cv.model(model)
\end{lstlisting}

In the code snippet, the function \code{cv.model} takes an argument
\code{x}, which we observed to always be a {\it linear model} (either ANOVA with \CLASS{aov, lm} or a simple linear regression model with \CLASS{lm}) which is part
of R's built-in statistics functionality.  Internally, linear models are
represented as lists with named elements, and one may access these named
elements using the \code{$} syntax, as seen on lines 2-4 in the snippet.  To
describe these kinds of values, we intended to include a struct type in our type
system, allowing users to specify name, type pairs as part of list type
definitions, though we found this to be problematic.

The inclusion of structs complicated our analysis by
polluting the data, illustrated on line 8 with the call
\code{data(sweetpotato)}.  It is quite common for R example code to use some
built-in data sets, such as \code{sweetpotato}, to show off some package
functionality with ``real data''.  When \typetracer is configured to detect
structs we saw a marked increase in misrepresentative types, as the types of
functions which use these data sets (such as \code{aov} in the code example) became polluted with the names from the example
sets.  We implemented a unification strategy for struct types, removing them
if they co-occurred with list types, and turning them into lists if two
co-occurring struct types shared no common names, but even so we found
ourselves with function signatures that were full of noise introduced by the
struct names.

We found that struct-like data was best described with classes, as illustrated in the code snippet, and many structs actually had a class.
Even though \code{x} is used like a struct, it in fact has a linear model class, which we believe is a more informative type.
